idlspec2d dataflow

### 2016-06-02 doc/algorithms/algorithms.tex
adding stuff to fluxdistortion
adding fluxcorr description

@@

The {\it flux distortion} vectors are parameterized in terms of magnitude
(i.e.~log-flux) that are achromatic with $x$, $y$, $x^2$, $y^2$, $xy$,
where those are linear coordinates XFOCAL,YFOCAL from the plugmap.
There are also chromatic terms that scale as
$\tilde \lambda = 1-(5070/\lambda)^2$,
since that function gives an equal effect between 3900 and 5070~\AA\
as between 5070 ang 9000~\AA.
There are also magnitude offsets as a function of spectrograph ID,
and a chromatic offset as a function of spectrograph ID.  The 14 parameters are:
\begin{equation}
\begin{split}
    F_\mathrm{new} &= F_\mathrm{orig}
        \left(1 + a_0 s_1 + a_1 s_2 \right) \mathrm{exp} \left( \right. \\
&            a_2 x + a_3 y + a_4 x y + a_5 x^2 + a_6 y^2 + \\
&            a_7 \tilde \lambda x + a_8 \tilde \lambda y + \\
&             a_9 \tilde \lambda s_1 + a_{10} \tilde \lambda s_2 +
             a_{11} \tilde \lambda^2 s_1 + a_{12} \tilde \lambda^2 s_2
        \left. \right)
\end{split}
\end{equation}
where $s_1$=1 if specid=1 else 0 and $s_2$=1 if specid=2 else 0.

>>>>

Individual exposures are initially flux calibrated with no constraint that the same object has the same flux across different exposures.  Emperical ``fluxcorr'' vectors are broadband corrections to bring the different exposures into alignment for each object prior to coaddition.  In DR13 and prior, these were implemented for each spectrum by minimizing
\begin{equation}
    \chi^2_i = \sum_\lambda {
        (f_{i\lambda} - f_{\mathrm{ref},\lambda} / a_{i\lambda} )^2 \over
        (\sigma_{i\lambda}^2 - \sigma_{\mathrm{ref},\lambda}^2 / a_{i\lambda}^2)
        }
\end{equation}
where $f_{i\lambda}$ is the flux of exposure $i$ at wavelength $\lambda$;
$f_{\mathrm{ref},\lambda}$ is the flux of the selected reference exposure;
and $a_{i\lambda}$ are low order Legendre polynomials.
%--- If you want more detail:
 The number of polynomial terms is dynamic, up to
 a maximum of 5 terms.  Higher order terms are added only if they improve
 the $\chi^2$ by 5 compared to one less term.
This approach is biased toward small $a_{i\lambda}$ since that inflates the denominator to reduce the $\chi^2$.

For DR14, we solve the fluxcorr vectors relative to a common weighted coadd $F_\lambda$ which is treated as noiseless compared to the individual exposures.
\begin{equation}
    \chi^2_i = \sum_\lambda {
        (f_{i\lambda} - F_\lambda / a_{i\lambda})^2 \over
        \sigma_{i\lambda}^2
        }
\end{equation}
We additionally include an emperically tuned prior that $a_{i\lambda} \sim 1$ to avoid large excursions in the solution for very low signal-to-noise data.

 The DR14 fluxcorr terms are Chebychev polynomials instead of Legendre for no
 good reason.  We actually solve for ((f - F a)/sigma)^ and then return 1/a .
 The prior is weighted by the data weights such that the relative strength of
 the prior vs. the data is approximately independent of S/N.  I'm not convinced
 that is the best thing to do but it worked better than having a fixed
 strength prior.

@@

The {\it flux distortion} vectors are parameterized in terms of magnitude
(i.e.~log-flux) that are achromatic with $x$, $y$, $x^2$, $y^2$, $xy$,
where those are linear coordinates XFOCAL,YFOCAL from the plugmap.
There are also chromatic terms that scale as
$\tilde \lambda = 1-(5070/\lambda)^2$,
since that function gives an equal effect between 3900 and 5070~\AA\
as between 5070 ang 9000~\AA.
There are also magnitude offsets as a function of spectrograph ID,
and a chromatic offset as a function of spectrograph ID.  The 13 parameters are:
\begin{equation}
\begin{split}
    F_\mathrm{new} &= F_\mathrm{orig}
        \left(1 + a_0 s_1 + a_1 s_2 \right) \mathrm{exp} \left( \right. \\
&            a_2 x + a_3 y + a_4 x y + a_5 x^2 + a_6 y^2 + \\
&            a_7 \tilde \lambda x + a_8 \tilde \lambda y + \\
&             a_9 \tilde \lambda s_1 + a_{10} \tilde \lambda s_2 +
             a_{11} \tilde \lambda^2 s_1 + a_{12} \tilde \lambda^2 s_2
        \left. \right)
\end{split}
\end{equation}
where $s_1$=1 if specid=1 else 0 and $s_2$=1 if specid=2 else 0.

Only \textsc{SPECTROPHOTO\_STD} or \textsc{REDDEN\_STD} objects are used to compute the flux distortion. The procedure minimizes differences between spectro-flux and photometry (CALIBFLUX). For now, only g, r and i-bands are used.
